---
title: "Assignment 02"
author: "Scott Stoltzman"
date: "5/12/2019"
output: html_document
---

Install package: 'nycflights123'

```{r}
library('nycflights13')
library('tidyverse')
```


### Answer the problems below in the spaces between `<ANSWER BELOW>` and `<ANSWER ABOVE>`

## Practice

1.  Why does this code not work?

    ```{r, error = TRUE}
    my_variable <- 10
    my_varıable
    ```
    
    Look carefully! (This may seem like an exercise in pointlessness, but
    training your brain to notice even the tiniest difference will pay off
    when programming.)
    
    
    `<ANSWER BELOW>`
    The call of my_varıable contains a dotless i character, in Unicode:
    (U+0131	ı	305	&inodot;	Latin Small Letter dotless I	0241)
    versus the dotted i (U+0069	i	105	Latin Small Letter I	0074).
    R returns an error since the dotless i variable was never instanciated.
    `<ANSWER ABOVE>`
    
    
1.  Press Alt + Shift + K. What happens? How can you get to the same place
    using the menus?


    `<ANSWER BELOW>`
    R opens the Keyboard Shortcuts Help (quick reference). From menu: Tools >  Keyboard Shortcuts Help
    `<ANSWER ABOVE>`



Find all flights that  

  1. Had an arrival delay of two or more hours  
  2. Flew to Houston (IAH or HOU)
  3. Were operated by United, American, or Delta
  4. Departed in summer (July, August, and September)
  5. Arrived more than two hours late, but didn’t leave late
  6. Were delayed by at least an hour, but made up over 30 minutes in flight
  7. Departed between midnight and 6am (inclusive)

```{r}
?flights
dat <- flights
filter(dat, arr_delay >= 120)
filter(dat, dest == "IAH" | dest == "HOU")
filter(dat, carrier == "UA" | carrier == "AA" | carrier == "DL")
filter(dat, month == 7 | month == 8 | month == 9)
filter(dat, ((arr_delay > 120) & (dep_delay <= 0)))
filter(dat, (dep_delay >= 60), (dep_delay > arr_delay + 30))
View(filter(dat, between(dep_time, 0000, 0600)))

```


Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?
```{r}
filter(dat, between(dep_time, 0000, 0600))
```


How many flights have a missing dep_time? What other variables are missing? What might these rows represent?
```{r}
sum(is.na(dat$dep_time)) # 8255 missing
sum(is.na(dat$year)) # 0 missing
sum(is.na(dat$month)) # 0 missing
sum(is.na(dat$day)) # 0 missing
sum(is.na(dat$arr_time)) #8713 missing
sum(is.na(dat$sched_dep_time)) # 0 missing
sum(is.na(dat$sched_arr_time)) # 0 missing
sum(is.na(dat$dep_delay)) #8255 missing
sum(is.na(dat$arr_delay)) # 9430 missing
sum(is.na(dat$carrier)) # 0 missing
sum(is.na(dat$flight)) # 0 missing
sum(is.na(dat$tailnum)) # 2512 missing
sum(is.na(dat$origin)) # 0 missing
sum(is.na(dat$dest)) # 0 missing
sum(is.na(dat$air_time)) #9430 missing
sum(is.na(dat$distance)) # 0 missing
sum(is.na(dat$hour)) # 0 missing
sum(is.na(dat$minute)) # 0 missing
sum(is.na(dat$time_hour)) # 0 missing
sum(is.na(dat$dep_time)) # 0 missing
## The flights with missing values never actually launched, so they probably were cancelled...
colSums(is.na(dat))
```

Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)
```{r}
NA ^ 0 # Results in 1. So, NA holds TRUE and not missing. 
NA | TRUE # Results in TRUE since (in the OR logical expression) one of the two is TRUE , and it's not missing.
FALSE & NA # The & returns TRUE if both are TRUE, so this results in FALSE. 
## If it can pass the logical expression, then it will return NA and not missing.
NA * 0 # Results in NA
```

